---
title: "effects.Rmd"
author: "sven panis"
date: "2024-05-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Here we will visualize the effects in various models.

# load the libraries that we will be using #

```{r load-pkg}
pkg <- c("standist", "tidyverse", "RColorBrewer", "patchwork", 
         "tidybayes", "bayesplot", "brms")

lapply(pkg, library, character.only = TRUE)
```

# Models build

```{r}
b0.general1 <- readRDS("models/b0.general1.rds")
b1.general1_prime <- readRDS("models/b1.general1.prime.rds") # model b
b1.general1_prime_prior <- readRDS("models/b1.general1_prime_prior.rds") # model c

b2.constant <- readRDS("models/b2.constant.rds")
b2.linear <- readRDS("models/b2.linear.rds")
b2.quad <- readRDS("models/b2.quad.rds")
b2.cubic <- readRDS("models/b2.cubic.rds")
b2.factor <- readRDS("models/b2.factor.rds")
b2.general <- readRDS("models/b2.general.rds")

b4_general_nointeractions <- readRDS("models/b4_general_nointeractions.rds")
b4_general_interactions2 <- readRDS("models/b4_general_interactions2.rds")
b4_general_interactions_nonlinear2 <- readRDS("models/b4_general_interactions_nonlinear2.rds")
b4_general_interactions_nonlinear_noprop3 <- readRDS("models/b4_general_interactions_nonlinear_noprop3.rds")

#b5_multilevel_1 <- readRDS("models/b5_multilevel_1.rds")
b5_multilevel_2f <- readRDS("models/b5_multilevel_2f.rds")


```

# Let's display fitted hazard and survivor functions for model c. 

First, we can create a table with the posterior means, ignoring uncertainty.

```{r}
tibble(time  = 6:15,
       alpha = fixef(b1.general1_prime_prior)[1:10, 1],
       beta_con  = fixef(b1.general1_prime_prior)[11, 1],
       beta_incon  = fixef(b1.general1_prime_prior)[12, 1]) %>% 
  mutate(lh0 = alpha, # logit-hazard
         lh_con = alpha + beta_con,
         lh_incon = alpha + beta_incon) %>% 
  mutate(h0 = inv_logit_scaled(lh0), # hazard
         h_con = inv_logit_scaled(lh_con),
         h_incon = inv_logit_scaled(lh_incon)) %>% 
  mutate(s0 = cumprod(1 - h0), # survival probabilities
         s_con = cumprod(1 - h_con),
         s_incon = cumprod(1 - h_incon)) %>% 
  # this just simplifies the output
  mutate_if(is.double, round, digits = 4)
```

To make a plot, we go beyond posterior means and reintroduce the uncertainty in the model.

For the disaggregated (ptb) data used to fit b1.general1_prime_prior model, here is how we might define the newdata, pump it through the model using fitted(), and wrangle.

```{r}
nd <-
  crossing(incon   = 0:1,
           con = 0:1,
           period = 6:15) %>% 
  filter(!(con == 1 & incon == 1)) %>%
  mutate(condition = rep(c("Neutral","Congruent","Incongruent"),each=10),
         condition = factor(condition)) %>%
  mutate(d6  = if_else(period == 6, 1, 0),
         d7  = if_else(period == 7, 1, 0),
         d8  = if_else(period == 8, 1, 0),
         d9  = if_else(period == 9, 1, 0),
         d10 = if_else(period == 10, 1, 0),
         d11 = if_else(period == 11, 1, 0),
         d12 = if_else(period == 12, 1, 0),
         d13 = if_else(period == 13, 1, 0),
         d14 = if_else(period == 14, 1, 0),
         d15 = if_else(period == 15, 1, 0))

f <-
  fitted(b1.general1_prime_prior,
         newdata = nd,
         scale = "linear") %>% 
  data.frame() %>% 
  bind_cols(nd) #%>% 
  #mutate(con = str_c("con = ", con),
  #       incon = str_c("incon = ", incon))

f
```

Now make a plot.

```{r}
# logit(hazard)
p1 <-
  f %>% 
  
  ggplot(aes(x = period, group = condition,
             fill = condition, color = condition)) +
  geom_ribbon(aes(ymin = Q2.5, ymax = Q97.5),
              linewidth = 0, alpha = 1/6) +
  geom_line(aes(y = Estimate)) +
  labs(subtitle = "fitted logit(hazard)",
       y = NULL) +
  coord_cartesian(ylim = c(-4, 1)) +
  theme(legend.background = element_rect(fill = "transparent"),
        legend.key = element_rect(color = "grey92"),
        legend.position = "inside",
        legend.position.inside = c(.1, .74))

# hazard
p2 <-
  f %>% 
  mutate_at(vars(Estimate, Q2.5, Q97.5), .funs = inv_logit_scaled) %>% 
  
  ggplot(aes(x = period, group = condition,
             fill = condition, color = condition)) +
  geom_ribbon(aes(ymin = Q2.5, ymax = Q97.5),
              linewidth = 0, alpha = 1/6) +
  geom_line(aes(y = Estimate)) +
  labs(subtitle = "fitted hazard",
       y = NULL) +
  coord_cartesian(ylim = c(0, .7)) +
  theme(legend.position = "none")

# survival
p3 <-
  f %>% 
  mutate_at(vars(Estimate, Q2.5, Q97.5), .funs = inv_logit_scaled) %>% 
  group_by(condition) %>% 
  mutate(s       = cumprod(1 - Estimate),
         s_lower = cumprod(1 - Q2.5),
         s_upper = cumprod(1 - Q97.5)) %>% 
  ungroup() %>%
  #select(period:d15, s:s_upper)
  
  ggplot(aes(x = period, group = condition,
             fill = condition, color = condition)) +
  geom_hline(yintercept = .5, color = "white") +
  #geom_segment(x = imls[1], xend = imls[1],
  #             y = -Inf, yend = .5,
  #             color = "white", linetype = 2) +
  #geom_segment(x = imls[2], xend = imls[2],
  #             y = -Inf, yend = .5,
  #             color = "white", linetype = 2) +
  geom_ribbon(aes(ymin = s_lower, ymax = s_upper),
              linewidth = 0, alpha = 1/6) +
  geom_line(aes(y = s)) + 
  scale_y_continuous(NULL, breaks = c(0, .5, 1)) +
  labs(subtitle = "fitted survival probability") +
  coord_cartesian(ylim = c(0, 1)) +
  theme(legend.position = "none")

p4 <-
  f %>% 
  mutate_at(vars(Estimate, Q2.5, Q97.5), .funs = exp) %>% 
  
  ggplot(aes(x = period, group = condition,
             fill = condition, color = condition)) +
  geom_ribbon(aes(ymin = Q2.5, ymax = Q97.5),
              linewidth = 0, alpha = 1/6) +
  geom_line(aes(y = Estimate)) +
  labs(subtitle = "fitted odds(hazard)",
       y = NULL) +
  coord_cartesian(ylim = c(0, 2)) +
  theme(legend.position = "none")
```

```{r}
(p1 / p2 / p3 / p4) &
  scale_fill_viridis_d(NULL, option = "A", end = .6) &
  scale_color_viridis_d(NULL, option = "A", end = .6) &
  scale_x_continuous("Time bin rank", breaks = 6:15, limits = c(6, 15), labels=c(6:15*40)) &
  theme(panel.grid = element_blank())
```


Plot marginal posteriors of model b:

```{r}
draws <- as_draws_df(b1.general1_prime)

draws %>% 
  pivot_longer(b_con:b_incon) %>% 
  
  ggplot(aes(x = value, y = name, fill = after_stat(x > 0))) +
  stat_slab() +
  scale_fill_manual(values = c("blue3", "red3")) +
  labs(x = "marginal posterior",
       y = NULL) +
  coord_cartesian(ylim = c(1.5, 2.5)) +
  theme(panel.grid = element_blank())
```

What proportion of posterior draws is zero or below?

```{r}
draws %>% 
  pivot_longer(b_con:b_incon) %>% 
  group_by(name) %>% 
  summarise(`percent zero or below` = 100 * mean(value <= 0))
```

95% Bayesian credible intervals 

```{r}
fixef(b1.general1_prime)

```

95% highest density intervals

```{r}
draws %>% 
  pivot_longer(b_con:b_incon) %>% 
  group_by(name) %>% 
  mean_hdi(value)
```

Transform posterior credible intervals to odds-ratio metric

```{r}
fixef(b1.general1_prime)[c("con", "incon"), -2] %>% exp()
```

```{r}
draws %>% 
  pivot_longer(b_con:b_incon) %>% 
  mutate(`odds ratio` = exp(value)) %>% 
  
  ggplot(aes(x = `odds ratio`, y = name)) +
  stat_interval(size = 5, .width = seq(from = .1, to = .9, by = .2)) +
  scale_color_grey("CI level:", start = .8, end = .2) +
  scale_x_continuous(breaks = 0:3) +
  ylab(NULL) +
  coord_cartesian(xlim = c(0, 3)) +
  theme(legend.position = "top",
        panel.grid = element_blank())
```


# Compare models (constant, linear, ... , general) in a graph

```{r}
make_fitted <- function(fit, scale, ...) {
  
  fitted(fit,
         newdata = nd,
         scale = scale,
         ...) %>% 
    data.frame() %>% 
    bind_cols(nd)
}
```

Plot fitted logit(hazard) - scale = linear

```{r}
nd <- tibble(period_factor = 6:15)

f <- make_fitted(b2.factor, scale = "linear") %>% rename(period = period_factor)

# this will simplify the `mutate()` code below
models <- c("constant", "linear", "quadratic", "cubic", "general")

nd <- tibble(period = 6:15)

f <-
  bind_rows(make_fitted(b2.constant, scale = "linear"),  # constant
            make_fitted(b2.linear, scale = "linear"),  # linear
            make_fitted(b2.quad, scale = "linear"),  # quadratic
            make_fitted(b2.cubic, scale = "linear"),  # cubic
            f) %>%                                   # general
  mutate(model = factor(rep(models, each = 10),
                        levels = models))

# what have we done?
glimpse(f)
```

```{r}
p1 <-
  f %>% 
  ggplot(aes(x = period, y = Estimate, color = model)) +
  geom_line() +
  scale_color_viridis_d(option = "A", direction = -1) +
  ylab("Fitted logit(hazard)") +
  coord_cartesian(ylim = c(-4, 1)) +
  theme(panel.grid = element_blank())
```

Plot fitted hazard

```{r}
nd <- tibble(period_factor = 6:15)

f <- make_fitted(b2.factor, scale = "response") %>% rename(period = period_factor)

nd <- tibble(period = 6:15)

f <-
  bind_rows(make_fitted(b2.constant, scale = "response"),  # constant
            make_fitted(b2.linear, scale = "response"),  # linear
            make_fitted(b2.quad, scale = "response"),  # quadratic
            make_fitted(b2.cubic, scale = "response"),  # cubic
            f) %>%                                     # general
  mutate(model = factor(rep(models, each = 10),
                        levels = models))
```

```{r}
p2 <-
  f %>% 
  filter(model %in% c("quadratic", "general")) %>% 
  
  ggplot(aes(x = period, y = Estimate, color = model)) +
  geom_line() +
  scale_color_viridis_d(option = "A", end = .5, direction = -1) +
  ylab("Fitted hazard") +
  coord_cartesian(ylim = c(0, .5)) +
  theme(legend.position = "none",
        panel.grid = element_blank())
```

Fitted survivor functions

```{r}
p3 <-
  f %>% 
  filter(model %in% c("quadratic", "general")) %>% 
  select(Estimate, period, model) %>% 
  # add the `new_rows` data
  #bind_rows(new_rows) %>%
  arrange(model, period) %>%
  group_by(model) %>% 
  # convert hazards to survival probabilities
  mutate(Estimate = cumprod(1 - Estimate)) %>%
  
  # plot!
  ggplot(aes(x = period, y = Estimate, color = model)) +
  geom_hline(yintercept = .5, color = "white") +
  geom_line() +
  scale_color_viridis_d(option = "A", end = .5, direction = -1) +
  scale_y_continuous("Fitted survival probability", breaks = c(0, .5, 1)) +
  coord_cartesian(ylim = c(0, 1)) +
  theme(legend.position = "none",
        panel.grid = element_blank())
```

```{r}
p1 + p2 + p3 + 
  plot_layout(guides = "collect") &
  scale_x_continuous("Time bin", breaks = 6:15, limits = c(6, 15))
```

# Plot posterior distributions of hazard ratios for con and incon, for trials 500,1000, and 1500 (models b4_general_nointeractions, b4_general_interactions2).

```{r}
as_draws_df(b4_general_interactions2) %>% # 8000 x 17 (b_d6, ... , )
  expand_grid(incon = 0:1,
              con  = 0:1,
              trial_c = c(-0.5,0,0.5)) %>% # 96000 rows
  filter(!(con ==1 & incon == 1)) %>% # 72000

  
  mutate(hr = exp(b_con * con + b_trial_c * trial_c  + `b_con:trial_c` * con * trial_c +
                  b_incon * incon + `b_incon:trial_c` * incon * trial_c),
         Condition       = rep(c("N","N","N","C","C","C","I","I","I"), 8000),
         trial_C = str_c("trial_c=",trial_c*1000+1000)) %>% 
  
  ggplot(aes(x = hr, y = 0)) +
  stat_halfeye(.width = .95, normalize = "panels") +
  scale_y_continuous(NULL, breaks = NULL) +
  theme(panel.grid = element_blank()) +
  facet_grid(Condition ~ trial_C)
```

Plot predicted cloglog(hazard)

```{r}
# define the `newdata`
nd <- tibble(period = 6:15) %>%
    expand_grid(incon = 0:1,
              con  = 0:1,
              trial_c = c(-0.5,0,0.5)) %>%
  filter(!(con ==1 & incon == 1)) %>%
  mutate(d6 = if_else(period == 6, 1, 0),
                           d7 = if_else(period == 7, 1, 0),
                           d8 = if_else(period == 8, 1, 0),
                           d9 = if_else(period == 9, 1, 0),
                           d10 = if_else(period == 10, 1, 0),
                           d11 = if_else(period == 11, 1, 0),
                           d12 = if_else(period == 12, 1, 0),
                           d13 = if_else(period == 13, 1, 0),
                           d14 = if_else(period == 14, 1, 0),
                           d15 = if_else(period == 15, 1, 0))
  

# use `fitted()` and wrangle
make_fitted(b4_general_interactions2, scale = "response") %>% 
  mutate(cond = rep(rep(c("N","C","I"),each=3),10)   ,
         trial    = factor(trial_c,levels=c(-0.5,0.0,0.5),labels=c("500","1000","1500"))) %>% 
  
  # plot!
  ggplot(aes(x = period, y = Estimate, ymin = Q2.5, ymax = Q97.5,
             fill = cond, color = cond)) +
  geom_ribbon(alpha = 1/5, linewidth = 0) +
  geom_line() +
  scale_fill_viridis_d(NULL, option = "A", end = .6, direction = -1) +
  scale_color_viridis_d(NULL, option = "A", end = .6, direction = -1) +
  scale_x_continuous("bin", breaks = 6:15, limits = c(6, 15)) +
  ylab("fitted cloglog(hazard)") +
  coord_cartesian(ylim = c(0, 1)) +
  theme(panel.grid = element_blank()) +
  facet_wrap(~ trial)
```

# model b5_multilevel_2f: Plot predicted cloglog(hazard) for single subject using fitted() in make_fitted().

```{r}
# define the `newdata`
nd <- tibble(pid = 1:6) %>%
  expand_grid(incon = 0:1,
                con  = 0:1,
                trial_c = c(-0.5,0,0.5),
                period_9 = -3:6) %>%
  filter(!(con ==1 & incon == 1)) %>%
  mutate(Iperiod_9E2 = period_9^2,
         Itrial_cE2 = trial_c^2)
  
# use `fitted()` and wrangle
make_fitted(b5_multilevel_2f, scale = "linear")%>% 
  mutate(cond   = rep(rep(c("N","C","I"),each=30),6),
         trial  = factor(trial_c,levels=c(-0.5,0.0,0.5),labels=c("500","1000","1500")),
         period = period_9+9) %>% 
  
  filter(pid==4) %>% # select pid here (1:6) manually
  
  # plot!
  ggplot(aes(x = period, y = Estimate, ymin = Q2.5, ymax = Q97.5,
             fill = cond, color = cond)) +
  geom_ribbon(alpha = 1/5, linewidth = 0) +
  geom_line() +
  scale_fill_viridis_d(NULL, option = "A", end = .6, direction = -1) +
  scale_color_viridis_d(NULL, option = "A", end = .6, direction = -1) +
  scale_x_continuous("bin", breaks = 6:15, limits = c(6, 15)) +
  ylab("fitted cloglog(hazard)") +
  coord_cartesian(ylim = c(-5, 1)) +
  theme(panel.grid = element_blank()) +
  facet_wrap(~ trial)
```

Same on hazard scale

```{r}
# define the `newdata`
nd <- tibble(pid = 1:6) %>%
  expand_grid(incon = 0:1,
                con  = 0:1,
                trial_c = c(-0.5,0,0.5),
                period_9 = -3:6) %>%
  filter(!(con ==1 & incon == 1)) %>%
  mutate(Iperiod_9E2 = period_9^2,
         Itrial_cE2 = trial_c^2)
  
# use `fitted()` and wrangle
make_fitted(b5_multilevel_2f, scale = "response")%>% 
  mutate(cond   = rep(rep(c("N","C","I"),each=30),6),
         trial  = factor(trial_c,levels=c(-0.5,0.0,0.5),labels=c("500","1000","1500")),
         period = period_9+9) %>% 
  
  filter(pid==5) %>% # select pid here (1:6) manually
  
  # plot!
  ggplot(aes(x = period, y = Estimate, ymin = Q2.5, ymax = Q97.5,
             fill = cond, color = cond)) +
  geom_ribbon(alpha = 1/5, linewidth = 0) +
  geom_line() +
  scale_fill_viridis_d(NULL, option = "A", end = .6, direction = -1) +
  scale_color_viridis_d(NULL, option = "A", end = .6, direction = -1) +
  scale_x_continuous("bin", breaks = 6:15, limits = c(6, 15)) +
  ylab("fitted hazard") +
  coord_cartesian(ylim = c(0, 1)) +
  theme(panel.grid = element_blank()) +
  facet_wrap(~ trial)
```

Plot Pareto k estimates

```{r}
loo(b5_multilevel_2f)$diagnostics %>% 
  data.frame() %>% 
  # attach the `id` values
  bind_cols(dataPS) %>% 
  mutate(id = 1:n()) %>%
  
  ggplot(aes(x = id, y = pareto_k)) +
  geom_point(alpha = 1/4) + 
  geom_text(data = . %>% filter(pareto_k > .2),
            aes(x = id + 100, label = id),
            size = 3, hjust = 0) +
  theme(panel.grid = element_blank())
```

https://cran.r-project.org/web/packages/tidybayes/vignettes/tidy-brms.html

get a list of raw model variable names so that we know what variables we can extract from the model:

```{r}
get_variables(b5_multilevel_2f)
```

A. Gather variable (e.g., r_pid) indices (e.g., 1, Intercept) into a separate column

```{r}
offsets <- b5_multilevel_2f %>%
                spread_draws(r_pid[subject,term]) # 528,000 obs. of 6 vars

offsets %>% head(10)

offsets %>% filter(term == "con") # 48,000 (6 pid x 8000 draws) x 6 columns (subject, term, r_pid, .chain, .iteration, .draw)
```

Point summaries (median, mean, mode) and intervals (qi or percentile intervals, hdi)

For example, calculate the median and 95% quantile interval of three variables

```{r}
b5_multilevel_2f %>%
  spread_draws(b_Intercept, b_con, b_incon) %>%
  median_qi()
```

Or the mean and 95% highest density interval of four variables

```{r}
b5_multilevel_2f %>%
  spread_draws(b_period_9, b_trial_c, sd_pid__Intercept, cor_pid__Intercept__period_9) %>%
  mean_hdi()
```

In separate rows:

```{r}
b5_multilevel_2f %>%
  gather_draws(b_period_9, b_trial_c, sd_pid__Intercept, cor_pid__Intercept__period_9) %>%
  mean_hdi()
```

For indexed model variables:

```{r}
rpid_info <- b5_multilevel_2f %>%
                spread_draws(r_pid[subject,term]) %>% 
                mean_qi() # 66 (6 pid x 11 terms) x 8 columns

print.data.frame(rpid_info)
```

get summary (including median absolute deviation or mad) and convergence diagnostics:

```{r}
b5_multilevel_2f %>%
                spread_draws(r_pid[subject,term]) %>% 
                summarise_draws()

b5_multilevel_2f %>%
  spread_draws(b_Intercept, b_con, b_incon) %>%
  summarise_draws()
```

Combining variables with different indices.
For example, the intercept for each subject (i.e., estimated cloglog-hazard in bin 9 for baseline condition = Neutral, trial 1000):

```{r}
b5_multilevel_2f %>% 
  spread_draws(b_Intercept, r_pid[subject,term]) %>%
  filter(term == "Intercept") %>% # 48,000 x 7
  mutate(subject_intercept = b_Intercept + r_pid) %>%
  median_qi(subject_intercept)
```

Plotting intervals with multiple probability levels

```{r}
b5_multilevel_2f %>% 
  spread_draws(b_Intercept, r_pid[subject,term]) %>%
  filter(term == "Intercept") %>% # 48,000 x 7
  mutate(subject_intercept = b_Intercept + r_pid) %>%
  median_qi(subject_intercept, .width = c(.95, .8, .5)) %>%
  
  ggplot(aes(y = subject, x = subject_intercept, xmin = .lower, xmax = .upper)) +
  geom_pointinterval() +
  theme_minimal()
```

Intervals with densities

```{r}
b5_multilevel_2f %>% 
  spread_draws(b_Intercept, r_pid[subject,term]) %>%
  filter(term == "Intercept") %>% # 48,000 x 7
  mutate(subject_intercept = b_Intercept + r_pid) %>%
  
  ggplot(aes(y = factor(subject), x = subject_intercept)) +
  ggdist::stat_halfeye(.width = c(.95, .8), point_interval = "median_qi")
```

Diff in subject-intercept between subjects 3 and 2?

```{r}
b5_multilevel_2f %>% 
  spread_draws(b_Intercept, r_pid[subject,term]) %>%
  filter(term == "Intercept") %>% # 48,000 x 7
  filter(subject == 2 | subject== 3) %>%
  pivot_wider(names_from = subject, values_from =  c(b_Intercept, r_pid)) %>%
  mutate(subject_intercept_diff23 = (b_Intercept_2 + r_pid_2) - (b_Intercept_3 + r_pid_3)) %>%
  
  ggplot(aes( x = subject_intercept_diff23)) +
  ggdist::stat_halfeye(.width = c(.95, .8), point_interval = "median_qi")
```


Diff in subject-intercept between subjects 5 and 6?

```{r}
b5_multilevel_2f %>% 
  spread_draws(b_Intercept, r_pid[subject,term]) %>%
  filter(term == "Intercept") %>% # 48,000 x 7
  filter(subject == 5 | subject== 6) %>%
  pivot_wider(names_from = subject, values_from =  c(b_Intercept, r_pid)) %>%
  mutate(subject_intercept_diff56 = (b_Intercept_5 + r_pid_5) - (b_Intercept_6 + r_pid_6)) %>%
  
  ggplot(aes( x = subject_intercept_diff56)) +
  ggdist::stat_halfeye(.width = c(.95, .8), point_interval = "median_qi")
```

## Compare functions

# as_draws_df() and spread_draws() and gather_draws()

```{r}
old <-as_draws_df(b5_multilevel_2f) %>%  # 8000 x 145 variables
  select(b_Intercept, b_con, b_incon) # 8000 x 3
  
new<-b5_multilevel_2f %>%
  spread_draws(b_Intercept, b_con, b_incon) %>% # 8000 x 6
  select(b_Intercept, b_con, b_incon) # 8000 x 3

new2 <-b5_multilevel_2f %>%
  gather_draws(b_Intercept, b_con, b_incon) %>% # 24000 x 5 / long format with .variable and .value, and automatic grouping
  pivot_wider(names_from = .variable, values_from = .value) %>% # 8000 x 6
  select(b_Intercept, b_con, b_incon) # 8000 x 3

  
all.equal(new,old) # TRUE
all.equal(new2,old) # TRUE
all.equal(new,new2) # TRUE


head(old)
head(new)
```

# brms: fitted(), posterior_epred(), posterior_linpred(), posterior_predict()
# tidybayes: add_epred_draws(), add_linpred_draws(), add_predicted_draws(), add_residual_draws()

For binomial cloglog-hazard model: epred -> hazard,  linpred -> cloglog-hazard, predict->1/0?

First, use fitted :

```{r}
# define the `newdata`
nd <- tibble(pid = 1:6) %>%
  expand_grid(incon = 0:1,
                con  = 0:1,
                trial_c = c(-0.5,0,0.5),
                period_9 = -3:6) %>%
  filter(!(con ==1 & incon == 1)) %>%
  mutate(Iperiod_9E2 = period_9^2,
         Itrial_cE2 = trial_c^2) # 540 x 7 (pid, incon, con, trial_c, period_9, Iperiod_9E2, Itrial_cE2)
  
# use `fitted()` and wrangle
 fitted(b5_multilevel_2f,
         newdata = nd,
         scale = "linear") %>% # 540 x 4 / cloglog-hazard predictions
    data.frame() %>% 
    bind_cols(nd)%>% 
    mutate(cond   = rep(rep(c("N","C","I"),each=30),6),
         trial  = factor(trial_c,levels=c(-0.5,0.0,0.5),labels=c("500","1000","1500")),
         period = period_9+9) %>% 
   
    filter(pid==4) %>% # select pid here (1:6) manually
  
  # plot!
  ggplot(aes(x = period, y = Estimate, ymin = Q2.5, ymax = Q97.5,
             fill = cond, color = cond)) +
  geom_ribbon(alpha = 1/5, linewidth = 0) +
  geom_line() +
  scale_fill_viridis_d(NULL, option = "A", end = .6, direction = -1) +
  scale_color_viridis_d(NULL, option = "A", end = .6, direction = -1) +
  scale_x_continuous("bin", breaks = 6:15, limits = c(6, 15)) +
  ylab("fitted cloglog(hazard)") +
  coord_cartesian(ylim = c(-5, 1)) +
  theme(panel.grid = element_blank()) +
  facet_wrap(~ trial)
```

Second, use fitted() for typical subject (re_formula = F)

```{r}
# define the `newdata`
nd <- tibble(incon = 0:1) %>%
  expand_grid(con  = 0:1,
                trial_c = c(-0.5,0,0.5),
                period_9 = -3:6) %>%
  filter(!(con ==1 & incon == 1)) %>%
  mutate(Iperiod_9E2 = period_9^2,
         Itrial_cE2 = trial_c^2) # 90 x 6 (incon, con, trial_c, period_9, Iperiod_9E2, Itrial_cE2)
  
# use `fitted()` and wrangle
 fitted(b5_multilevel_2f,
         newdata = nd,
         scale = "linear",
        re_formula = NA) %>% # 90 x 4 / cloglog-hazard predictions
    data.frame() %>% 
    bind_cols(nd)%>% 
    mutate(cond   = rep(rep(c("N","C","I"),each=30),1),
         trial  = factor(trial_c,levels=c(-0.5,0.0,0.5),labels=c("500","1000","1500")),
         period = period_9+9) %>% 
    #filter(pid==4) %>% # select pid here (1:6) manually
  
  # plot!
  ggplot(aes(x = period, y = Estimate, ymin = Q2.5, ymax = Q97.5,
             fill = cond, color = cond)) +
  geom_ribbon(alpha = 1/5, linewidth = 0) +
  geom_line() +
  scale_fill_viridis_d(NULL, option = "A", end = .6, direction = -1) +
  scale_color_viridis_d(NULL, option = "A", end = .6, direction = -1) +
  scale_x_continuous("bin", breaks = 6:15, limits = c(6, 15)) +
  ylab("fitted cloglog(hazard)") +
  coord_cartesian(ylim = c(-5, 1)) +
  theme(panel.grid = element_blank()) +
  facet_wrap(~ trial)
```

Typical subject on hazard scale:

```{r}
# define the `newdata`
nd <- tibble(incon = 0:1) %>%
  expand_grid(con  = 0:1,
                trial_c = c(-0.5,0,0.5),
                period_9 = -3:6) %>%
  filter(!(con ==1 & incon == 1)) %>%
  mutate(Iperiod_9E2 = period_9^2,
         Itrial_cE2 = trial_c^2) # 90 x 6 (incon, con, trial_c, period_9, Iperiod_9E2, Itrial_cE2)
  
# use `fitted()` and wrangle
 fitted(b5_multilevel_2f,
         newdata = nd,
         scale = "response",
        re_formula = NA) %>% # 90 x 4 / cloglog-hazard predictions
    data.frame() %>% 
    bind_cols(nd)%>% 
    mutate(cond   = rep(rep(c("N","C","I"),each=30),1),
         trial  = factor(trial_c,levels=c(-0.5,0.0,0.5),labels=c("500","1000","1500")),
         period = period_9+9) %>% 
    #filter(pid==4) %>% # select pid here (1:6) manually
  
  # plot!
  ggplot(aes(x = period, y = Estimate, ymin = Q2.5, ymax = Q97.5,
             fill = cond, color = cond)) +
  geom_ribbon(alpha = 1/5, linewidth = 0) +
  geom_line() +
  scale_fill_viridis_d(NULL, option = "A", end = .6, direction = -1) +
  scale_color_viridis_d(NULL, option = "A", end = .6, direction = -1) +
  scale_x_continuous("bin", breaks = 6:15, limits = c(6, 15)) +
  ylab("fitted hazard") +
  coord_cartesian(ylim = c(0, 1)) +
  theme(panel.grid = element_blank()) +
  facet_wrap(~ trial)
```

Third, use add_epred_draws() and compare.

```{r}
# define the `newdata`
nd <- tibble(pid = 1:6) %>%
  expand_grid(incon = 0:1,
                con  = 0:1,
                trial_c = c(-0.5,0,0.5),
                period_9 = -3:6) %>%
  filter(!(con ==1 & incon == 1)) %>%
  mutate(Iperiod_9E2 = period_9^2,
         Itrial_cE2 = trial_c^2) # 540 x 7 (pid, incon, con, trial_c, period_9, Iperiod_9E2, Itrial_cE2)
  
# use `add_epred_draws()` and wrangle
 add_epred_draws(b5_multilevel_2f,
         newdata = nd,
         re_formula = NULL) %>% # 4,320,000 (=540x8000) x 12 / hazard predictions !!
    #data.frame() %>% 
    #bind_cols(nd)%>% 
   ungroup() %>%
  #  mutate(cond   = rep(rep(c("N","C","I"),each=30*8000),6),
  #       trial  = factor(trial_c,levels=c(-0.5,0.0,0.5),labels=c("500","1000","1500")),
  #       period = period_9+9) %>% 
  #  filter(pid==4) %>% # select pid here (1:6) manually
   group_by(.row)%>%
    mutate(epred = mean_qi(.epred)) %>%
   select(.epred, epred) %>%
   slice(1)

# compare to `fitted()` 
 fitted(b5_multilevel_2f,
         newdata = nd,
         scale = "response") # same values for epred and Estimate!!!
 
   
```


Third, use add_linpred_draws().

```{r}
# define the `newdata`
nd <- tibble(pid = 1:6) %>%
  expand_grid(incon = 0:1,
                con  = 0:1,
                trial_c = c(-0.5,0,0.5),
                period_9 = -3:6) %>%
  filter(!(con ==1 & incon == 1)) %>%
  mutate(Iperiod_9E2 = period_9^2,
         Itrial_cE2 = trial_c^2) # 540 x 7 (pid, incon, con, trial_c, period_9, Iperiod_9E2, Itrial_cE2)

# use `add_epred_draws()` and wrangle
 add_linpred_draws(b5_multilevel_2f,
         newdata = nd,
         re_formula = NULL)  # 4,320,000 (=540x8000) x 12 / cloglog-hazard .linpred !!
```
 

Fourth, use add_predicted_draws().

```{r}
# define the `newdata`
nd <- tibble(pid = 1:6) %>%
  expand_grid(incon = 0:1,
                con  = 0:1,
                trial_c = c(-0.5,0,0.5),
                period_9 = -3:6) %>%
  filter(!(con ==1 & incon == 1)) %>%
  mutate(Iperiod_9E2 = period_9^2,
         Itrial_cE2 = trial_c^2) # 540 x 7 (pid, incon, con, trial_c, period_9, Iperiod_9E2, Itrial_cE2)
  
# use `add_epred_draws()` and wrangle
 add_predicted_draws(b5_multilevel_2f,
         newdata = nd,
         re_formula = NULL) # 4,320,000 (=540x8000) x 12 / 0/1 .prediction !!
``` 






## generate fake data for Knot study, and fit max random effects model




