---
title: "Exercise1"
author: "sven panis"
date: "2024-04-07"
output: html_document
editor_options: 
  chunk_output_type: console
---

Create fake RT + acc data for 2 conditions (normal, exponential; about 30% errors each) of 1 subject, set-up life table, and plot h(t), S(t), and ca(t) per condition, together with median RT, and save figure.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages, include=F}
library(tidyverse) # data wrangling functions
library(patchwork) # combine figures
```

```{r options-tibble}
old <- options(
  pillar.sigfig = 4
)
```

Create a fake RT + accuracy data set for 1 subject.

```{r create-fake-data, echo=T}
set.seed(1274)
Trials_perCondition = 100

data <- tibble(
  pid   = 1, # person ID
  trial = sample(1:(2*Trials_perCondition)),
  condition = c(rep(1,Trials_perCondition), rep(2,Trials_perCondition)),
  RT = c(rnorm(Trials_perCondition, 800, 200), rexp(Trials_perCondition, 1/600)) %>% round(digits=2),
  acc = rep(sample(c(0,1), size=Trials_perCondition,replace=T, prob=c(.3,.7)),2)) %>% 
        arrange(trial) %>%
        mutate(RT = ifelse(RT<0,0.00000001,RT),
               condition=factor(condition,levels=c(1,2)))

print.data.frame(data, max=50)
#print(data, n=10) # ignores the round(digits=2) function
```

Create right-censored observations and discrete RT variable.

```{r preprocess}
cutoff <- 1100 # censoring time in ms
bin_size <- 100

data2 <- data %>%
         mutate(right_censored = 0,
                RT2 = ifelse(RT > cutoff, cutoff, RT) %>% round(digits=2),
                right_censored = ifelse(RT2 == cutoff,1,right_censored),
                dRT = ceiling(RT2/bin_size),
                cens_time = cutoff, bin_width = bin_size) # save for plotting

print.data.frame(data2,max=100)
```

Create person-trial-period data set.

```{r ptb-dataset}
data_ptb <-
  data2 %>% 
  uncount(weights = dRT) %>% # duplicate rows according to a weighting variable (or expression).
  group_by(trial) %>% 
  mutate(period = 1:n()) %>% 
  mutate(event = if_else(period == max(period) & right_censored == 0, 1, 0)) %>% 
  ungroup()

print.data.frame(data_ptb, max=200)
```

Set-up life table.

```{r lifetable}
data_lt <-
  data_ptb %>% 
  mutate(event = str_c("event", event)) %>%
  group_by(condition,period) %>% 
  count(event) %>% 
  ungroup() %>% 
  pivot_wider(names_from = event,
              values_from = n) %>% 
  mutate(ev0 = ifelse(is.na(event0),0,event0), # replace NA with 0
         ev1 = ifelse(is.na(event1),0,event1),
         risk_set = ev0 + ev1) %>%
  mutate(hazard = (ev1 / risk_set) %>% round(digits = 3)) %>% # estimate hazard
  mutate(se_haz = sqrt((hazard * (1 - hazard)) / risk_set) %>% round(digits=3)) %>% # se hazard
  group_by(condition) %>%
  mutate(survival = cumprod(1-hazard) %>% round(digits = 4), # estimate survival
         term     = cumsum(hazard / (risk_set * (1 - hazard))) %>% round(digits = 6), 
         se_surv = (survival * sqrt(term)) %>% round(digits = 3)) %>% # Greenwood's (1926) approximation
  ungroup()

print.data.frame(data_lt, max =200)
```

Calculate conditional accuracy.

```{r ca, message=F}
ca <- data2 %>%
  filter(right_censored==0) %>%
  group_by(condition,dRT,cens_time,bin_width) %>%
  summarize(ca = mean(acc) %>% round(digits=2),
            n = n()) %>%
  ungroup() %>%
  mutate(period = dRT,
         se_ca = sqrt((ca * (1-ca)) / n) %>% round(digits=3)) %>%
  select(-dRT)

print.data.frame(ca, max = 100)
```

Add conditional accuracy, and optionally the row for time 0 in the lifetables.

```{r full-lifetable}
lt_full <- data_lt %>%
    left_join(ca, join_by(condition,period))

row0_cond1 <- tibble(condition=factor(1),period=0,event0 = NA,event1 =NA, ev0 = NA,ev1 = NA,
                     risk_set = Trials_perCondition,hazard = NA,se_haz = NA,survival = 1,
                     term = NA,se_surv = 0,ca = NA,n = NA,se_ca = NA)
row0_cond2 <- tibble(condition=factor(2),period=0.0,event0 = NA,event1 = NA,ev0 = NA,ev1 = NA,
                     risk_set = Trials_perCondition,hazard = NA,se_haz = NA,survival = 1,
                     term = NA,se_surv = 0,ca = NA,n = NA,se_ca = NA)
                   
lt_full <- bind_rows(row0_cond1, row0_cond2, lt_full) %>% arrange(condition, period)
  
print.data.frame(lt_full,max=200)
```

Extract median RT for each condition, based on survivor functions.

```{r median-rt}
# extract rows around percentile 50 of the survivor function
above_pct50 <- lt_full %>% 
      group_by(condition) %>%
      filter(survival > .5) %>% 
      slice(n()) # take last row
below_pct50 <- lt_full %>% 
      group_by(condition) %>%
      filter(survival < .5) %>% 
      slice(1) # take first row

period_above <- pull(above_pct50, period)

# pull survivor function values
survival_above <- pull(above_pct50, survival)
survival_below <- pull(below_pct50, survival)

# estimate median by interpolation
median_period <- period_above+((survival_above-.5)/(survival_above-survival_below))*((period_above+1)-period_above)

# transform to median RT
median_RT <- median_period*bin_size
median_RT
```

Plot hazard, survival, S(t).50, and ca per condition.

```{r plotting, warning=F}
p1 <-lt_full %>%
  ggplot(aes(x=period, color=condition, group=condition)) +
  geom_line(aes(y=hazard)) +

  geom_point(aes(y=hazard), size=1) + labs(color="Condition") +
  geom_linerange(aes(ymin=hazard-se_haz, ymax=hazard+se_haz), show.legend = F) +
  scale_x_continuous(breaks = c(0,1:(cutoff/bin_size)), labels=c(0,1:(cutoff/bin_size)*bin_size),
                     limits = c(0,cutoff/bin_size)) +
  scale_y_continuous(limits = c(0,1)) +
  labs(x="", y="h(t)") +
  theme(legend.background = element_rect(fill = "transparent"),
        panel.grid = element_blank(),
        legend.position = "top")

p2 <-lt_full %>%
  ggplot(aes(x=period, color=condition, group=condition)) +
  geom_line(aes(y=survival), show.legend = F) +
  geom_point(aes(y=survival), size=1, show.legend = F) +
  geom_linerange(aes(ymin=survival-se_surv, ymax=survival+se_surv), show.legend = F) +
  # add medians
  geom_path(aes(x=period, y=survival, color=factor(condition)),
            data = tibble(period= c(median_period[1], median_period[1],
                                    median_period[2], median_period[2]),
                          survival = c(.5, 0, .5, 0),
                          condition = c(1,1,2,2)), 
            linetype = 3, show.legend = F) +
  scale_x_continuous(breaks = c(0,1:(cutoff/bin_size)), labels=c(0,1:(cutoff/bin_size)*bin_size),
                     limits=c(0,cutoff/bin_size)) +
  scale_y_continuous(limits = c(0,1)) +
  labs(x="", y="S(t)",
       colour="Condition") +
  theme(panel.grid = element_blank())

p3 <-lt_full %>%
  ggplot(aes(x=period, color=condition, group=condition)) +
  geom_line(aes(y=ca), show.legend = F) +
  geom_point(aes(y=ca), size=1, show.legend = F) +
  geom_linerange(aes(ymin=ca-se_ca, ymax=ca+se_ca), show.legend = F) +
  scale_x_continuous(breaks = c(0,1:(cutoff/bin_size)), labels=c(0,1:(cutoff/bin_size)*bin_size),
                     limits=c(0,cutoff/bin_size)) +
  scale_y_continuous(limits = c(0,1)) +
  labs(x="Time bin t's endpoint (ms)", y="ca(t)",
       colour="Condition") +
  theme(panel.grid = element_blank())

p1/p2/p3

ggsave("figures/example1.png", width = 8, height = 5, dpi = 600)
```

```{r set-back-tibble-options}
options(old)
```
