---
title: "Exercise3"
author: "sven panis"
date: "2024-04-17"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages, include=F}
library(tidyverse) # data wrangling
```

In Exercise 3 we will set up life tables for a real data set - experiment 1 from Panis & Schmidt (2016) - and compare our value to theirs to confirm the code in Exercise 2 is working correctly.

First, load the functions defined in Exercise2.Rmd.

Second, load the original datafile from Panis and Schmidt (2016), wrangle and reduce the datafile to 3 conditions, and apply the functions.
The original datafile contains RT and accuracy data for 12 conditions (3 prime types x 4 mask types) of 6 subjects. The censoring time used was 600 ms, and the bin width 40 ms.
Five trials without response info were deleted.

```{r process-data-file}
datafile.orig <- read_delim(file="data/DataExp1_6subjects.txt", delim="\t",col_names=T)

head(datafile.orig)
# remove practice trials
datafile.orig <- datafile.orig %>% filter(practice==0) # 9240 trials

# remove unnecessary columns
datafile.orig <- datafile.orig %>% select(-c(ss,practice,rd,form,task,primeduration,
                                             blankduration,maskduration,targetduration,
                                             void,target_dir,corresp,resperr))
# remove 5 trials with no response info.
cutoff <- 600
sel = which(datafile.orig$resp==0 & datafile.orig$rt <= cutoff)
sel
datafile.orig <- datafile.orig[-c(770,1425,1428,1511,1512),]
summary(datafile.orig)

# set required column names: pid, trial, condition, RT, and acc
colnames(datafile.orig) <- c("pid","bl","tr","mask","condition","resp","acc","RT","trial")

# keep 3 conditions: no mask (0), and no prime (0),consistent(1),inconsistent prime(2) 
datafile.orig <- datafile.orig %>% filter(mask == 0) # 2757 rows left
datafile.orig <- datafile.orig %>% mutate(condition = condition + 1, 
                                          condition = factor(condition, levels = c(1,2,3)))

# apply the functions
data_nested <- datafile.orig %>% group_nest(pid)

data_final <- data_nested %>% 
                     mutate(censored  = map(data, define_censoring_dRT, 600, 40)) %>% # user input: censoring time, and bin width
                     mutate(ptb_data  = map(censored, setup_person_trial_bin)) %>%           # create person-trial-bin dataset
                     mutate(lifetable = map(ptb_data, setup_lifetable)) %>%                  # create life tables without ca
                     mutate(condacc   = map(censored, calc_conditionalaccuracy)) %>%         # calculate ca
                     mutate(lifetable_ca = map2(lifetable, condacc, join_lifetable_ca)) %>%  # create life tables with ca
                     mutate(plot      = map2(.x = lifetable_ca, .y = pid, plot_eha))         # create plots of info in life tables
```

Third, load the corresponding life tables from Panis and Schmidt (2016), and compare the values.

```{r compare-results}
load("data/NCE_Panis_final_LifeTables.RData")

# compare lifetables for condition 1 and subject 1:
LifeTableSubject[[1]][[1]]

print.data.frame(data_final$lifetable_ca[[1]] %>% filter(condition == 1))

# conditon 2 and subject 1:

LifeTableSubject[[1]][[2]]

print.data.frame(data_final$lifetable_ca[[1]] %>% filter(condition == 2))
```

Save the plots for each subject.

```{r save-figs, eval=F}
# save as pdf
map2(paste0("figures/figure_for_subject", data_final$pid, "_PanisSchmidt.pdf"), data_final$plot, ggsave)
```

Now, assume all trials are independent, to fit bayesian hazard models without multilevel structure. Make 1 plot for all data.
This requires a single pid, and a trial column that identifies each unique trial.

```{r}
summary(datafile.orig)
# create new pid and trial variables
datafile.orig_ind <- datafile.orig %>% mutate(oldtrial=trial,
                                              oldpid=pid,
                                              trial = 1:n(),
                                              pid = 1)
summary(datafile.orig_ind)
```


```{r}
data_nested_ind <- datafile.orig_ind %>% group_nest(pid)

data_final_ind <- data_nested_ind %>% 
                     mutate(censored  = map(data, define_censoring_dRT, 600, 40)) %>% #3 user input: censoring time, and bin width
                     mutate(ptb_data  = map(censored, setup_person_trial_bin)) %>%           #4 create person-trial-bin dataset
                     mutate(lifetable = map(ptb_data, setup_lifetable)) %>%                  # create life tables without ca
                     mutate(condacc   = map(censored, calc_conditionalaccuracy)) %>%         # calculate ca
                     mutate(lifetable_ca = map2(lifetable, condacc, join_lifetable_ca)) %>%  # create life tables with ca
                     mutate(plot      = map2(.x = lifetable_ca, .y = pid, plot_eha))         # create plots of info in life tables
```

```{r}
# view person-trial-bin data set 
print.data.frame(data_final_ind[[4]][[1]], max=400) 

# view lifetables for each condition 
print.data.frame(data_final_ind[[5]][[1]], max=600)
```


```{r save-plot-ind}
# save as pdf
map2(paste0("figures/figure_for_ind", data_final_ind$pid, ".pdf"), data_final_ind$plot, ggsave)
```

Extract file for modeling hazard without multilevel structure.

```{r file-hazard-modeling}
input_modeling_hazard_ind <- data_final_ind %>% 
                               select(pid, ptb_data) %>%  unnest(ptb_data) %>%
                               select(pid,trial,condition,period,event)

# save in folder data
write_csv(input_modeling_hazard_ind, file="data/inputfile_hazard_modeling_ind.csv")
```

Now, assume all trials are independent, and only 1 condition (to compare with model b0.general1).
This requires a single pid, a single condition, and a trial column that identifies each unique trial.

```{r}
summary(datafile.orig)
# create new pid and trial variables
datafile.orig_ind_1cond <- datafile.orig %>% mutate(oldtrial=trial,
                                              oldpid=pid,
                                              oldcond = condition,
                                              trial = 1:n(),
                                              pid = 1,
                                              condition = 1)
summary(datafile.orig_ind_1cond)
```

```{r}
data_nested_ind_1cond <- datafile.orig_ind_1cond %>% group_nest(pid)

data_final_ind_1cond <- data_nested_ind_1cond %>% 
                     mutate(censored  = map(data, define_censoring_dRT, 600, 40)) %>% #3 user input: censoring time, and bin width
                     mutate(ptb_data  = map(censored, setup_person_trial_bin)) %>%    #4 create person-trial-bin dataset
                     mutate(lifetable = map(ptb_data, setup_lifetable)) 
```

```{r}
# view lifetables for single condition 
print.data.frame(data_final_ind_1cond[[5]][[1]], max=600)

# the MLE hazard estimates in bins 6-15 are almost identical to the estimates of "fixef(b0.general1) %>% inv_logit()", as expected.
```
