---
title: "Exercise2"
author: "sven panis"
date: "2024-04-10"
output: html_document
editor_options: 
  chunk_output_type: console
---
Extend Exercise1 to multiple subjects using functional programming:
Create fake RT + acc data for x conditions of y subjects, set-up life table, 
and plot h(t), S(t), and ca(t) for each condition per subject, together with median RT.
Save figures, and save one file for hazard modeling, and one for ca(t) modeling, and print a nice life table.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages, include=F}
library(tidyverse) # data wrangling
library(patchwork) # combine figures
library(pracma)    # to Reshape() matrices
#(flextable) # create tables 
```

Create a fake RT (normal) + accuracy (about 70%) data set with x conditions for y subjects, and z trials per condition.

```{r create-fake-data, echo=T}
set.seed(1276)
Tr_pCond = 100 # trials per condition
N_sub = 4    # number of subjects
N_cond = 3   # number of conditions per subject
N_total = N_sub*N_cond*Tr_pCond

data <- tibble(
  pid   = rep(1:N_sub,each=N_cond*Tr_pCond),
  trial = as.integer(Reshape( replicate(N_sub,sample(1:(N_cond*Tr_pCond))),N_total,1)),
  condition = rep(rep(1:N_cond,each=Tr_pCond), N_sub),
  RT = as.double(Reshape(replicate(N_sub*N_cond,rnorm(Tr_pCond, 600, 200)),N_total,1)) %>% round(digits=2),
  acc = as.double(Reshape(replicate(N_sub*N_cond,(sample(c(0,1), size=Tr_pCond,replace=T,            
                                                  prob=c(.3,.7)))),N_total,1))) %>%
      arrange(pid,trial) %>%
      mutate(RT = ifelse(RT<0,0.001,RT), # instead of filtering to keep all rows
             condition=factor(condition,levels=c(1:N_cond)))

print.data.frame(data, max=50)
```

Nest data within subjects.

```{r nest-subjects}
data_nested <- data %>% group_nest(pid)
```

Now define and apply various functions using map and map2, to set-up life tables and create plots showing the descriptive functions of discrete time.

Checkout these websites:
https://modern-rstats.eu/functional-programming.html#functional-programming-and-plotting
https://dcl-prog.stanford.edu/purrr-basics.html#extra-arguments

```{r func-define-censoring-drt}
define_censoring_dRT <- function(df, timeout, bin_width){
  if(!(timeout %% bin_width == 0)){
    return("The censoring time must be a multiple of the bin width!")
  }
  df %>% mutate(right_censored = 0,
                RT2 = ifelse(RT > timeout, timeout, RT) %>% round(digits=2),
                right_censored = ifelse(RT2 == timeout,1,right_censored),
                dRT = ceiling(RT2/bin_width),
                cens_time = timeout, bin_width = bin_width)
}
```

```{r func-setup-ptb}
setup_person_trial_bin <- function(df){
  df %>% uncount(weights = dRT) %>% 
         group_by(trial) %>% 
         mutate(period = 1:n()) %>% 
         mutate(event = if_else(period == max(period) & right_censored == 0, 1, 0)) %>% 
         ungroup()
}
```

```{r func-setup-lifetable}
setup_lifetable <- function(ptb){
  ptb %>% mutate(event = str_c("event", event)) %>%
          group_by(condition,period) %>% 
          count(event) %>% 
          ungroup() %>% 
          pivot_wider(names_from = event,
                      values_from = n) %>% 
          mutate(event0 = ifelse(is.na(event0),0,event0), # replace NA with 0
                 event1 = ifelse(is.na(event1),0,event1),
                 risk_set = event0 + event1) %>% # define the risk set
          mutate(hazard = (event1 / risk_set) %>% round(digits = 3)) %>% # calculate hazard estimate
          mutate(se_haz = sqrt((hazard * (1 - hazard)) / risk_set) %>% round(digits = 3)) %>% # se hazard
          group_by(condition) %>%
          mutate(survival = (cumprod(1-hazard)) %>% round(digits = 4), # calculate survival estimate
                 term     = cumsum(hazard / (risk_set * (1 - hazard))) %>% round(digits = 3),
                 se_surv  = (survival * sqrt(term)) %>% round(digits = 3)  ) %>% # Greenwood's (1926) approximation
          ungroup() 
}
```

```{r func-conditional-accuracy}
calc_conditionalaccuracy <- function(df){
  df %>% filter(right_censored==0) %>%
         group_by(condition,dRT,cens_time,bin_width) %>%
         summarize(ca = mean(acc) %>% round(digits = 2),
                   n = n(),
                   .groups = 'drop') %>%
         ungroup() %>%
         mutate(period = dRT,
                se_ca = sqrt((ca * (1-ca)) / n) %>% round(digits = 3)) %>%
         select(-dRT)
}
```

```{r func-join}
join_lifetable_ca <- function(df1,df2){df1 %>% left_join(df2, join_by(condition,period))}
```

```{r func-extract-median}
extract_median_period <- function(df){
  above_pct50 <- df %>% 
      group_by(condition) %>%
      filter(survival > .5) %>% 
      slice(n()) # take last row
  below_pct50 <- df %>% 
      group_by(condition) %>%
      filter(survival < .5) %>% 
      slice(1) # take first row
  # pull period above
  period_above <- pull(above_pct50, period)
  # pull survivor function values
  survival_above <- pull(above_pct50, survival)
  survival_below <- pull(below_pct50, survival)
  # estimate median by interpolation
  median_period <- period_above+((survival_above-.5)/(survival_above-survival_below))*((period_above+1)-period_above)
}
```

```{r func-make-plots}
plot_eha <- function(df,subj){
  library(patchwork)
  cutoff <- df %>% pull(cens_time) %>% max(na.rm=T)
  binsize <- df %>% pull(bin_width) %>% max(na.rm=T)
  median_period <- extract_median_period(df)
  n_conditions <- nlevels(df$condition)
  data_median <- c()
  for(i in 1:n_conditions){
    data_median <- append(data_median, c(median_period[i], median_period[i]))
  }
  data_medians <- tibble(period= data_median,
                         survival = rep(c(.5, 0),n_conditions),
                         condition = rep(1:n_conditions, each=2))
  
p1 <- df %>% ggplot(aes(x=period, color=condition, group=condition)) +
  geom_line(aes(y=hazard)) +
  geom_point(aes(y=hazard), size=1) + labs(color="Condition") +
  geom_linerange(aes(ymin=hazard-se_haz, ymax=hazard+se_haz), show.legend = F) +
  scale_x_continuous(breaks = c(0,1:(cutoff/binsize)), labels=c(0,1:(cutoff/binsize)*binsize),
                     limits = c(0,cutoff/binsize)) +
  scale_y_continuous(limits = c(0,1)) +
  labs(x="", y="h(t)", title = paste("Subject ", subj)) +
  theme(legend.background = element_rect(fill = "transparent"),
        panel.grid = element_blank(),
        legend.position = "top")

p2 <-df %>%
  ggplot(aes(x=period, color=condition, group=condition)) +
  geom_line(aes(y=survival), show.legend = F) +
  geom_point(aes(y=survival), size=1, show.legend = F) +
  geom_linerange(aes(ymin=survival-se_surv, ymax=survival+se_surv), show.legend = F) +
  # add medians
  geom_path(aes(x=period, y=survival, color=factor(condition)),
            data = data_medians, 
            linetype = 3, show.legend = F) +
  scale_x_continuous(breaks = c(0,1:(cutoff/binsize)), labels=c(0,1:(cutoff/binsize)*binsize),
                     limits=c(0,cutoff/binsize)) +
  scale_y_continuous(limits = c(0,1)) +
  labs(x="", y="S(t)",
       colour="Condition") +
  theme(panel.grid = element_blank())

p3 <-df %>%
  ggplot(aes(x=period, color=condition, group=condition)) +
  geom_line(aes(y=ca), show.legend = F) +
  geom_point(aes(y=ca), size=1, show.legend = F) +
  geom_linerange(aes(ymin=ca-se_ca, ymax=ca+se_ca), show.legend = F) +
  scale_x_continuous(breaks = c(0,1:(cutoff/binsize)), labels=c(0,1:(cutoff/binsize)*binsize),
                     limits=c(0,cutoff/binsize)) +
  scale_y_continuous(limits = c(0,1)) +
  labs(x="Time bin t's endpoint (ms)", y="ca(t)",
       colour="Condition") +
  theme(panel.grid = element_blank())

p1/p2/p3
}
```

Apply the functions using map and map2 to create the life tables and plots.

```{r apply-functions}
data_final <- data_nested %>% 
                     mutate(censored  = map(data, define_censoring_dRT, 800, 100)) %>%       # user input: censoring time, and bin width
                     mutate(ptb_data  = map(censored, setup_person_trial_bin)) %>%           # create person-trial-bin dataset
                     mutate(lifetable = map(ptb_data, setup_lifetable)) %>%                  # create life tables without ca
                     mutate(condacc   = map(censored, calc_conditionalaccuracy)) %>%         # calculate ca
                     mutate(lifetable_ca = map2(lifetable, condacc, join_lifetable_ca)) %>%  # create life tables with ca
                     mutate(plot      = map2(.x = lifetable_ca, .y = pid, plot_eha))         # create plots of info in life tables
```

Now you can extract info as required.

```{r extract-info}
# view original data for subject 1:
print.data.frame(data_final[[2]][[1]], max=60)

# view person-trial-bin data set for subject 1:
print.data.frame(data_final[[4]][[1]], max=160) 

# view lifetables for each condition for subject 1
print.data.frame(data_final[[7]][[1]], max=300)
```

And you can save each plot.

```{r save-plots}
# save as pdf
map2(paste0("figures/figure_for_subject", data_final$pid, ".pdf"), data_final$plot, ggsave)

# save as png
map2(paste0("figures/figure_for_subject", data_final$pid, ".png"), data_final$plot, ggsave, width = 8, height = 5, dpi = 600)
```

Make a nice life table for condition 1 of subject 1, and include row for time zero.

```{r make-table}
tbl <- data_final[[7]][[1]] %>% 
  filter(condition==1) %>% 
  select(period,risk_set,hazard,se_haz,survival,se_surv,ca, se_ca)

knitr::kable(as.matrix(tbl),"simple", align = "lrcccccc", digits=c(0,0,2,4,2,4,2,4))
# TODO: add row for time zero
```

Extract file for modeling hazard.

```{r file-hazard-modeling}
input_modeling_hazard <- data_final %>% 
                        select(pid, ptb_data) %>%  unnest(ptb_data) %>%
                        select(pid,trial,condition,period,event)

# save in folder data
write_csv(input_modeling_hazard, file="data/inputfile_hazard_modeling.csv")
```

Extract file for modeling ca(t).

```{r file-ca-modeling}
input_ca_modeling <- data_final %>%
                    select(pid, censored) %>% unnest(censored) %>%
                    filter(right_censored == 0) %>% # only keep uncensored trials
                    select(pid, trial, condition, acc, dRT)

# save in folder data
write_csv(input_ca_modeling, file="data/inputfile_ca_modeling.csv")
```

Extract median RTs.

```{r extract-medians}
LL <-map(data_final$lifetable_ca, extract_median_period)
binsize <- data_final$censored[[1]] %>% pull(bin_width) %>% max(na.rm=T)
modify_depth(LL,1,~.x*binsize)
```
