---
title: "Exercise1"
author: "sven panis"
date: "2024-04-07"
output: html_document
---

Create fake RT + acc data for 2 conditions (normal, exponential; about 30% errors each) of 1 subject, set-up life table, and plot h(t), S(t), and ca(t) per condition, together with median RT.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r loadPck, include=F}
library(tidyverse)
library(patchwork)
```

Create a fake RT + accuracy data set for 1 subject.

```{r createFakeData, echo=F}
set.seed(1274)
Trials_perCondition = 1000

data <- tibble(
  pid   = 1,
  trial = sample(1:(2*Trials_perCondition)),
  condition = c(rep(1,Trials_perCondition), rep(2,Trials_perCondition)),
  RT = c(rnorm(Trials_perCondition, 800, 200), rexp(Trials_perCondition, 1/600)),
  acc = rep(sample(c(0,1), size=Trials_perCondition,replace=T, prob=c(.3,.7)),2)) %>% 
        arrange(trial) %>%
        mutate(RT = ifelse(RT<0,0.00000001,RT),
               condition=factor(condition,levels=c(1,2)))

print.data.frame(data)
```

Create right-censored observations and discrete RT variable.

```{r preprocess}
cutoff <- 1000 # censoring time in ms
bin_size <- 100

data2 <- data %>%
         mutate(censor = 1,
                RT2 = ifelse(RT > cutoff, cutoff, RT),
                censor = ifelse(RT2 == cutoff,0,censor),
                dRT = ceiling(RT2/bin_size))

print.data.frame(data2)
```


Create person-trial-period data set.

```{r ptbDataSet}
data_ptb <-
  data2 %>% 
  uncount(weights = dRT) %>% # duplicate rows according to a weighting variable (or expression).
  group_by(trial) %>% 
  mutate(period = 1:n()) %>% 
  mutate(event = if_else(period == max(period) & censor == 1, 1, 0)) %>% 
  select(-censor) %>% 
  ungroup()

#print.data.frame(data_ptb)
```

Set-up life table.

```{r lifetable}
data_lt <-
  data_ptb %>% 
  mutate(event = str_c("event = ", event)) %>%
  group_by(condition,period) %>% 
  count(event) %>% 
  ungroup() %>% 
  pivot_wider(names_from = event,
              values_from = n) %>% 
  mutate(ev1 = ifelse(is.na(`event = 1`),0,`event = 1`), # replace NA with 0
         ev0 = ifelse(is.na(`event = 0`),0,`event = 0`),
         RS = ev0 + ev1) %>%
  mutate(hazard = (ev1 / RS) %>% round(digits = 3)) %>% # estimate hazard
  mutate(se_haz = sqrt((hazard * (1 - hazard)) / RS) %>% round(digits=3)) %>% # se hazard
  group_by(condition) %>%
  mutate(survival = cumprod(1-hazard) %>% round(digits = 3), # estimate survival
         term     = cumsum(hazard / (RS * (1 - hazard))) %>% round(digits = 6), 
         se_surv = (survival * sqrt(term)) %>% round(digits = 7)) %>% # Greenwood's (1926) approximation
  ungroup()

print.data.frame(data_lt)
```

Calculate conditional accuracy.

```{r ca}
ca <- data2 %>%
  filter(censor==1) %>%
  group_by(condition,dRT) %>%
  summarize(ca = mean(acc),
            n = n()) %>%
  ungroup() %>%
  mutate(period = dRT,
         se_ca = sqrt((ca * (1-ca)) / n)) %>%
  select(-dRT)

ca
```

Add conditional accuracy...

```{r fulllifetable}
lt_full <- data_lt %>%
    left_join(ca, join_by(condition,period))

print.data.frame(lt_full)
```

Extract median RTs based on survivor functions.

```{r medianRTs}

```


Plot hazard, survival, S(t).50, and ca per condition.

```{r plotting}
p1 <-data_lt %>%
  ggplot(aes(x=period, color=factor(condition), group=condition)) +
  geom_line(aes(y=hazard)) +
  geom_point(aes(y=hazard), size=1) + labs(color="Condition") +
  geom_linerange(aes(ymin=hazard-se_haz, ymax=hazard+se_haz), show.legend = F)+
  #scale_color_viridis_d(NULL, option = "A", end=.55) +
  scale_x_continuous(breaks = 1:(cutoff/bin_size), labels=c(1:(cutoff/bin_size)*bin_size)) +
  scale_y_continuous(limits = c(0,1)) +
  labs(x="Time bin endpoint (ms)",
       y="hazard probability") +
  theme(legend.background = element_rect(fill = "transparent"),
        panel.grid = element_blank(),
        legend.position = "top")

p2 <-data_lt %>%
  ggplot(aes(x=period, color=factor(condition), group=condition)) +
  geom_line(aes(y=survival), show.legend = F) +
  geom_point(aes(y=survival), size=1, show.legend = F) +
  geom_linerange(aes(ymin=survival-se_surv, ymax=survival+se_surv), show.legend = F)+
  #scale_color_viridis_d(NULL, option = "A", end=.55) +
  scale_x_continuous(breaks = 1:(cutoff/bin_size), labels=c(1:(cutoff/bin_size)*bin_size)) +
  scale_y_continuous(limits = c(0,1)) +
  labs(x="Time bin endpoint (ms)",
       y="survival probability",
       colour="Condition") +
  theme(legend.background = element_rect(fill = "transparent"),
        panel.grid = element_blank())

p3 <-lt_full %>%
  ggplot(aes(x=period, color=factor(condition), group=condition)) +
  geom_line(aes(y=ca), show.legend = F) +
  geom_point(aes(y=ca), size=1, show.legend = F) +
  geom_linerange(aes(ymin=ca-se_ca, ymax=ca+se_ca), show.legend = F)+
  #scale_color_viridis_d(NULL, option = "A", end=.55) +
  scale_x_continuous(breaks = 1:(cutoff/bin_size), labels=c(1:(cutoff/bin_size)*bin_size)) +
  scale_y_continuous(limits = c(0,1)) +
  labs(x="Time bin endpoint (ms)",
       y="conditional accuracy",
       colour="Condition") +
  theme(legend.background = element_rect(fill = "transparent"),
        panel.grid = element_blank())



p1/p2/p3

ggsave("example1.png", width = 8, height = 5, dpi = 600)

```




